### 2023-11-01 journal/conclusion.tex
restored details about vector implementation in the concluding remarks

We believe the approach used here to implement a lock-free queue 
could be applied to obtain other lock-free
data structures with a polylogarithmic step complexity.
For example, we can easily adapt our routines to implement a  vector data structure that stores a sequence and
provides three operations: \opa{Append}{e} to add an element \var{e} to the end of the sequence,
\opa{Get}{i} to read the \var{i}th element in the sequence, and
\opa{Index}{e} to compute the position of element \var{e} in the sequence.
%\here{Can omit rest of this paragraph to save space}
%An \opa{append}{e} is implemented like \opa{Enqueue}{e} in $O(\log p)$ steps.  
%A \opa{get}{i} is similar to \op{GetEnqueue}, taking $O(\log n + \log^2p)$ steps when the vector has $n$ elements.  
%An \opa{index}{e} is similar to \op{IndexDequeue} (except operating on enqueues instead of dequeues) and would take $O(\log p)$ steps.
\here{doublecheck all preceding claims for accuracy.}
Building on the work described in this paper, Asbell and Ruppert \cite{AR23} have
designed a doubly-ended queue with polylogarithmic amortized step
complexity, which also yields a stack as a special case.  
This required a substantially different representation of the data stored in the ordering tree.
Whether the ordering tree could also be used to obtain a priority queue with polylogarithmic step complexity
remains an open question.

>>>>

We believe the approach used here to implement a lock-free queue 
could be applied to obtain other lock-free
data structures with a polylogarithmic step complexity.
For example, we can easily adapt the routines of the implementation in Section \ref{DescriptQ} 
to implement a  vector data structure that stores a sequence and
provides three operations: \opa{Append}{e} to add an element \var{e} to the end of the sequence,
\opa{Get}{i} to read the \var{i}th element in the sequence, and
\opa{Index}{e} to compute the position of element \var{e} in the sequence.
Only the \op{Append} operations need to be propagated to the root of the ordering tree
since the other two operations do not affect the state of the object.
An \opa{Append}{e} is implemented like \opa{Enqueue}{e} in $O(\log p)$ steps.  
A \opa{Get}{i} is similar to \op{GetEnqueue}, taking $O(\log n + \log^2p)$ steps when the vector has $n$ elements.  
An \op{Index} is similar to \op{IndexDequeue} (except operating on enqueues instead of dequeues) and would take $O(\log p)$ steps if the argument is a pointer to the leaf block that contains the element~$e$.

Building on the work described in this paper, Asbell and Ruppert \cite{AR23} have
designed a doubly-ended queue with polylogarithmic amortized step
complexity, which also yields a stack as a special case.  
This required a substantially different representation of the data stored in the ordering tree.
Whether the ordering tree could also be used to obtain a priority queue with polylogarithmic step complexity
remains an open question.

### 2023-11-01 journal/related.tex
restored details about vector implementation in the concluding remarks

Fatourou and Kallimanis \cite{FK14} used their own universal construction based on fetch\&add and LL/SC instructions
to implement a queue, but its step complexity is also $\Omega(p)$.
\here{Double check this.}

>>>>

Fatourou and Kallimanis \cite{FK14} used their own universal construction based on fetch\&add and LL/SC instructions
to implement a queue, but its step complexity is also $\Omega(p)$ when words are of reasonable size.
